A simple model to describe how outflows and rotation shape the Lyman-$\alpha$ emission line in galaxies

### 2017-03-28 paper/paper.tex
intro

Distant galaxies are key to understand early evolutionary stages of
our Universe.
However, those galaxies are challenging to detect.
Physical conditions in those galaxies allows the emergence of
Lyman-$\alpha$ line emission at $1216$ \AA \cite{PartrigePeebles}.
Galaxies detected through its Lyman-$\alpha$ emission receive the name of
spectra at  and named Lyman Alpha Emitters (LAEs).

Currently LAEs are commonly targetted in wide area galaxy surveys.
They have been effectively used to study galaxy evolution,
cosmology and the thermal history of the Universe.
This has been able through the study of their spatial distribution and
their global line shape.
The recent release of MUSE has caused an switch in the kind of studies
that can be performed on LAEs.

Thanks to MUSE's ability to spatially resolve a galaxy, we are now able
to infer detailed kinematic maps for nearby galaxies.
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

On the ISM's features that plays an important role in shaping the \lya
is HI kinematics. 
In a static HI medium the \lya line has  two equal and symmetric peaks
around the natural \lya wavelength and zero intensity at the line's
center.
If the ISM has an outflow velocity profile, the line becomes
asymmetrical with a more pronounced read peak.
If the galaxy rotates, the line shows different amounts of Doppler
shifts that would also modify the profile.
gas has a bulk velocity, the shape of the \lya profile changes. We
explore these effects in this paper by proposing a new model for a
LAE. The model consists of a spherical distribution of H atoms
undergoing a solid body rotation and radial expansion (outflows). \\

>>>>

Distant galaxies are key to understand early evolutionary stages of
our Universe.
Physical conditions in those galaxies allows the emergence of
Lyman-$\alpha$ line emission at $1216$ \AA \cite{PartrigePeebles}.
Galaxies detected through its Lyman-$\alpha$ emission receive the name of
spectra at  and named Lyman Alpha Emitters (LAEs).

Currently LAEs are commonly targetted in wide area galaxy surveys.
They have been effectively used to study galaxy evolution,
cosmology and the thermal history of the Universe.
This has been able through the study of their spatial distribution and
the shape of the \lya emission line.

Recent imptovements in instrumentation have revolutionized the kind of
studies that can be performed on LAEs.  
It is now possible to infer detailed kinematic maps for nearby galaxies.
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

On the ISM's features that plays an important role in shaping the \lya
is HI kinematics. 
In a static HI medium the \lya line has  two equal and symmetric peaks
around the natural \lya wavelength and zero intensity at the line's
center.
For an outflowing ISM, the line becomes
asymmetrical with a more pronounced read peak.
If the galaxy rotates, the line shows different amounts of Doppler
shifts modifying the overall line profile \cite{Garavito14}.

In this paper we present for the first time a study of the joint
effects of galaxy outflows and rotation.
We study a simplified geometrical configuration corresponding to an
spherical gas cloud with symmetrical radial outflows and a rotation
profile corresponding to a solid body.
